---
documentclass: jss
author:
  - name: Łukasz Chrostowski
    affiliation: Adam Mickiewicz University
    affiliation2: Pearson
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{lukchr@st.amu.edu.pl}
    url: https://posit.co
  - name: Piotr Chlebicki
    orcid: 0009-0006-4867-7434
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Stockholm University \AND
    email: \email{piotr.chlebicki@math.su.se}
  - name: Maciej Beręsewicz
    orcid: 0000-0002-8281-4301
    address: |
      | Department of Statistics,
      | Institute of Informatics and Electronic Economy,
      | Poznań University of Economics and Business
    affiliation: |
      | Poznań University of Economics and Business
    affiliation2: 'Statistical Office in Poznań'
    # To add another line, use \AND at the end of the previous one as above
    email: \email{maciej.beresewicz@ue.poznan.pl}
    url: https://posit.co
title:
  formatted: "\\pkg{nonprobsvy} -- An R package for modern methods for non-probability surveys"
  # If you use tex in the formatted title, also supply version without
  plain:     "nonprobsvy -- An R package for modern methods for non-probability surveys"
  # For running headers, if needed
  short:     "\\pkg{nonprobsvy} for non-probability surveys"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{R}"]
  plain:     [keywords, not capitalized, R]
preamble: >
  \usepackage{amsmath}
  \newcommand{\bX}{\boldsymbol{X}}
  \newcommand{\bx}{\boldsymbol{x}}
  \newcommand{\bY}{\boldsymbol{Y}}
  \newcommand{\by}{\boldsymbol{y}}
  \newcommand{\bh}{\boldsymbol{h}}
  \newcommand{\bH}{\boldsymbol{H}}
  \newcommand{\ba}{\boldsymbol{a}}
  \newcommand{\bp}{\boldsymbol{p}}
  \newcommand{\bA}{\boldsymbol{A}}
  \newcommand{\bw}{\boldsymbol{w}}
  \newcommand{\bd}{\boldsymbol{d}}
  \newcommand{\bZ}{\boldsymbol{Z}}
  \newcommand{\bz}{\boldsymbol{z}}
  \newcommand{\bv}{\boldsymbol{v}}
  \newcommand{\bu}{\boldsymbol{u}}
  \newcommand{\bU}{\boldsymbol{U}}
  \newcommand{\bQ}{\boldsymbol{Q}}
  \newcommand{\bG}{\boldsymbol{G}}
  \newcommand{\HT}{\text{\rm HT}}
  \newcommand{\bbeta}{\boldsymbol{\beta}}
  \newcommand{\balpha}{\boldsymbol{\alpha}}
  \newcommand{\btau}{\boldsymbol{\tau}}
  \newcommand{\bgamma}{\boldsymbol{\gamma}}
  \newcommand{\btheta}{\boldsymbol{\theta}}
  \newcommand{\blambda}{\boldsymbol{\lambda}}
  \newcommand{\bPhi}{\boldsymbol{\Phi}}
  \newcommand{\bEta}{\boldsymbol{\eta}}
  \newcommand{\bZero}{\boldsymbol{0}}
  \newcommand{\colvec}{\operatorname{colvec}}
  \newcommand{\logit}{\operatorname{logit}}
  \newcommand{\Exp}{\operatorname{Exp}}
  \newcommand{\Ber}{\operatorname{Bernoulli}}
  \newcommand{\Uni}{\operatorname{Uniform}}
output: rticles::jss_article
bibliography: references.bib
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

# Introduction

This template demonstrates some of the basic LaTeX that you need to know to create a JSS article.

## Code formatting

In general, don't use Markdown, but use the more precise LaTeX commands instead:

* \proglang{Java}
* \pkg{plyr}

One exception is inline code, which can be written inside a pair of backticks (i.e., using the Markdown syntax).

If you want to use LaTeX commands in headers, you need to provide a `short-title` attribute. You can also provide a custom identifier if necessary. See the header of Section \ref{r-code} for example.

# Methods for non-probability samples \proglang{R} code {short-title="R code" #r-code}

## Basic setup

Let $U=\{1,..., N\}$ denote the target population consisting of $N$ labelled units. Each unit $i$ has an associated vector of auxiliary variables $\bx_{i}$ (a realisation of the random vector $\bX_{i}$ in the super-population) and the study variable $y_{i}$ (a realisation of the random variable $Y_{i}$ in the super-population). Let $\{ (y_i, \bx_i), i \in S_A\}$ be a dataset of a non-probability sample of size $n_A$ and let  $\{\left(\bx_i, \pi_{i}\right), i \in S_B\}$ be a dataset of a probability sample of size $n_B$, where only information about variables $\bX$ and inclusion probabilities $\pi$ (which in the super population model are also considered to be random variables) are available. Let $\delta$ be an indicator of inclusion into non-probability sample. Each unit in the sample $S_B$ has been assigned a~design-based weight given by $d_i = 1/\pi_i$. The setting is summarised in Table .. 

The goal is to estimate a~finite population mean $\displaystyle\mu_{y}=\frac{1}{N}\sum_{i=1}^{N} y_{i}$ of the target variable $Y$. As values of $y_{i}$ are not observed in the probability sample, it cannot be used to estimate the target quantity. Instead, one could try combining the non-probability and probability samples to estimate $\mu_{y}$. In this paper we do not consider modifications for the possibly occurring overlap.


## Mass Imputation estimators

## Inverse Probability Weighting estimators

## Doubly Robust estimators

Can be inserted in regular R markdown blocks.

```{r}
x <- 1:10
x
```

## Features specific to \pkg{rticles} {short-title="Features specific to rticles"}

* Adding short titles to section headers is a feature specific to \pkg{rticles} (implemented via a Pandoc Lua filter). This feature is currently not supported by Pandoc and we will update this template if [it is officially supported in the future](https://github.com/jgm/pandoc/issues/4409).
* Using the `\AND` syntax in the `author` field to add authors on a new line. This is a specific to the `rticles::jss_article` format.

# Package contents and implementation

# Practical examples

# Summary

